% Challenge in wireless communication
Due to the convenience of deployment and utility, wireless devices and network
have become the dominant tools for telecommunication. However, the spectrum resource
is limited in natural and has to be approved for accessibility, there is a huge 
challenge for individual devices and infrastructure to efficiently exploit the 
licensed frequency. In 2009, the FCC has approved the use of broadband services 
in the white space frequency of UHF TV bands, which were formerly exclusively 
licensed to television broadcasters. These white space bands are now available 
for unlicensed public use, offerring new opportunities for new design of device 
and network with better performance in throughput and economy cost. while many 
metropolitan areas sought to deploy city-wide WiFi networks, the densest urban 
areas were not able to broadly leverage the technology for large-scale Internet 
access.  Ultimately, the small spatial separation required for effective 802.11 
links in these areas resulted in prohibitively large up-front costs. The far 
greater range of these lower white space carrier frequencies are especially 
critical in rural areas, where high levels of aggregation could dramatically 
lower the cost of deployment and is in direct contrast to dense urban areas in 
which networks are built to maximize spatial reuse. Through both academy and 
industry are eagerly looking forward to the application of these white space 
bands, more solid theoritic and pratical work need to be done of this topic. 
In this work, we investigate multiband adaptation in both link communication 
and network deployment to answer part of the questions. Furthermore, we design 
measurement driven algorithms to implement multiband spectrum agility in both 
the wireless link and wireless networks.
% List previous papers
% VNC
First, wireless channel selection is the foundation and one of the key parts 
of further wireless applications. We build a measurement driven algorithm 
to discover the best channel in vehicular multiband environment. We leverage 
knowledge of in-situ operation across frequency bands with real-time measurements 
of the activity level to select the the band with the highest throughput. To 
do so, we perform a number of experiments in typical vehicular topologies. 
With two models based on machine learning algorithms and an in-situ training 
set, we predict the throughput based on: ({\it i.}) prior performance for 
similar context information ({\it e.g.}, SNR, GPS, relative speed, and link 
distance), and ({\it ii.}) real-time activity level and relative channel quality 
per band. In the field, we show that training on a repeatable route with these 
machine learning techniques can yield vast performance improvements from prior 
schemes. 
% Winmee
Second, leveraging a broad range of spectrum across diverse population densities 
becomes a critical issue for the deployment of data networks with WiFi and white 
space bands. To address this issue we propose a measurement-driven band selection 
framework, Multiband Access Point Estimation (MAPE). Then we apply the framework 
with the data measured the spectrum utility in the Dallas-Fort Worth metropolitan 
and surrounding areas to show the white space band benefit in network deployment. 
In particular, we study the white space and WiFi bands with in-field spectrum 
utility measurements, revealing the number of access points required for an area 
with channels in multiple bands. In doing so, we find that networks with white space 
bands reduce the number of access points by up to 1650\% in sparse rural areas over 
similar WiFi-only solutions. In more populated rural areas and sparse urban areas, 
we find an access point reduction of 660\% and 412\%, respectively.  However, due 
to the heavy use of white space bands in dense urban areas, the cost reductions 
invert (an increase in required access points of 6\%).  Finally, we numerically 
analyze band combinations in typical rural and urban areas and show the critical 
factor that leads to cost reduction: considering the same total number of channels, 
as more channels are available in the white space bands, less access points are 
required for a given area.
% Whiteteris
Furthermore, we compare the heterogeneous access points with multiband spectrum
in variety configurations to discover the best set of spectrum combinations 
for an access point. In particular, we model the heterogeneous white space and 
WiFi access tier deployment problem and propose a relaxed ILP to get the lower 
bound of the amount of access point under resource limitations and a heuristic 
approach of the problem. Then, we map the problem as a Bin packing problem and 
propose Multiband Hetegeneous Access Point Deployment(MHAPD) method to address
the problem. In doing so, we discuss the benefit of white space bands 
in reducing the number of acess points and provide heuristic solution for 
wireless network deployment. The numberic results shows that MHAPD gains 260\% 
against the linear hexagon deployment method.  We further analysis the performance 
of hetergenous access point performance in variation of population density. 
The numberic results shows that hetergenous access point could improve the budget 
saving upto 323\%. 
% WM
Then, to quantify the spectrum utilization and far greater range of lower carrier 
frequencies on multihop access networks, 
%we consider how these white 
%space bands can be leveraged in large-scale wireless mesh network deployments 
%with in-field measured channel utilization. In particular, 
we present an integer 
linear programming model to leverage diverse propagation characteristics and 
the  channel occupancy of white space and WiFi bands to deploy optimal WhiteMesh 
networks. Since such problem is known to be NP-hard, we design a measurement driven 
heuristic algorithm, Band-based Path Selection (BPS), which we show approaches 
the performance of the optimal solution with reduced complexity.  We additionally 
compare the performance of BPS against two well-known multi-channel, multi-radio 
deployment algorithms across a range of scenarios spanning those typical for 
rural areas to urban areas. In doing so, we achieve up to 160\% traffic achieved 
gateways gain of these existing multi-channel, multi-radio algorithms, which are 
agnostic to diverse propagation characteristics across bands.  Moreover, we show 
that, with similar channel resources and bandwidth, the joint use of WiFi and 
white space bands can achieve a served user demand of 170\% that of mesh networks 
with only WiFi bands or white space bands, respectively. Further, through the 
result, we leverage the channel occupancy and spacing impacts on white mesh
network and study the general rules of band selection.

% Future work
Lastly, we propose to study new solution for multiband application in large 
scale network deployment. Graph theory is a efficient way to model and solve 
the deployment problem. The wireless multiband network deployment is formulated
as a graph theory problem to locate the gateway nodes. In our previous work,
we focus on solve the channel assignment and access tier mesh nodes deployment
in multiband scenarios. With the propose work of graph solution of gateway nodes
location, a complete process of multiband wireless network solution could be 
built. Also we are going to discuss beamforming in wireless network deployment. 
Beamforming provides potential gains for reducing the deployment cost and improving
the fairness in wireless network deployment. We are going to exploit the great
spacial agility from beamforming for future wireless networks.

