\chapter{Introduction} \label{ch:introduction}

Wireless network is the most convenient method to get service for users.
The demand of wireless device users inspires vendors to provide faster,
stable service in their networks. Because of the prevalence of individual
devices, vehicular wireless devices and wireless sensors, wireless data
traffic is expected to increasde a thousand-fold over the next decade~\cite{metis},
greatly motivating the improvement of link communication and network
deployment.

However, unlike wired networks, wireless devices have to share crowd
and unstable channels in limited frequencies. Especially in vehicular
environment, the situation is even worse due to the moving. 
Fortunately, drivers and passengers around the world could utilize a 
wide array of vehicular applications ranging from real-time traffic 
monitoring and safety applications to various infotainment applications.
However, the continuous use of such applications is limited due to the
challenge of transmitting over highly-dynamic vehicular wireless channels. 
In such networks, the increasing availability of different 
frequency bands with correspondingly diverse propagation characteristics could allow flexibility and 
robustness of vehicular links. Even with spectral flexibility, links are extremely tenuous, 
demanding instantaneous decisions to remain connected, motivating an algorithm that
can find the appropriate frequency band quickly and according to the current environmental context.

Obviously, more frequencies could improve the performance for both link and network communication.
The FCC has approved the use of broadband services in the white spaces of 
UHF TV bands, which were formerly exclusively licensed to television broadcasters.
These white space bands are now available for unlicensed public use, enabling the
deployment of wireless access networks across a broad range of scenarios from 
sparse rural areas (one of the key applications identified by the FCC) to dense urban 
areas~\cite{carlson}. The white space bands operate in available channels from 
54-806 MHz, having a far greater propagation range than WiFi bands for similar
transmission power~\cite{balanis2012antenna}. 
In WiFi and white space heterogeneous wireless network, the service area degree of 
an access point depends on the capacity of radios, the propagation range and the 
demands of the serving area. The scant frequencies of radios, the propagation distinctive
and the demands diversity of population distribution bring the variation of an access point
service area. These issues are substantial to designing an optimal network deployment 
and provide potential commercial wireless services to clients in any location.
Thus, the new opportunities created by white spaces motivate the following 
questions for wireless Internet carriers, which have yet to be addressed: 
(i) To what degree can white space bands reduce the network deployment cost of
sparsely populated rural areas as opposed to comparable WiFi-only solutions? and 
(ii) Where along the continuum of user population densities do the white
space bands no longer offer cost savings for wireless network deployments?
and To what degree can hetergeneous access points benefit the dense 
population areas and sparsely populated rural areas?


White space is not only benefit the access tier network, but also benefit the 
backhual tier networks. About a decade ago, numerous cities solicited proposals from
network carriers for exclusive rights to deploy city-wide WiFi,
spanning hundreds of square miles.  
While the vast majority of the resulting awarded contracts used
a wireless mesh topology, initial field tests revealed that the 
actual WiFi propagation could not achieve the proposed mesh node
spacing. As a result, many network carriers opted to pay millions of 
dollars in penalties rather than face the exponentially-increasing
deployment costs (e.g., Houston~\cite{cnet_aug07} and 
Philadelphia~\cite{arstechnica_may08}). Thus, while a few mesh 
networks have been deployed in certain communities~\cite{CRSK06,google_imc08},
wireless mesh networks have largely been unsuccessful in achieving 
the scale of what was once anticipated~\cite{taps}.
Around the same time, the digital TV transition created more
spectrum for use with data networks~\cite{fccwhitespace}. These white 
space bands operate in available channels from 54-806 MHz, having
increased propagation characteristics as compared to 
WiFi~\cite{balanis2012antenna}. Hence, the FCC has identified rural
areas as a key application for white space networks since the reduced
population from major metropolitan areas allows a greater service area
per backhaul device without saturating wireless capacity. Naturally, 
the question arises for these rural communities as well as more dense 
urban settings: how can the emerging white space bands improve 
large-scale mesh network deployments?  While much work has been done 
on deploying multihop wireless networks with multiple channels and 
radios, the differences in propagation have not be exploited in their 
models~\cite{raniwala2004centralized,tang2005interference, si2010overview}, 
which could be the fundamental issue for the success of mesh 
networks going forward.

Beamforming is recognized as an prospective technology to 
further improve the performance of wireless networks, 
through increasing both time reuse and spatial reuse. However,
people still have the question: in what degree beamforming could 
add capacity and reduce the cost for the network. This is the 
fundamental issue which could inspire new technology utilization
in the wireless networks.

\section{Challenges}
% Challenges

For wireless networks, both the clients and the vendors require
better service and low cost. However, due to the nature and policy, 
wireless channels suffer a lot in providing efficient and stable service for 
either link communication or network deployments. Spectrum 
adaptation offers new solution for these issues. However, the
diversity of spectrum and network complexity present great
challenges in designing multiband communication systems, 
for both link and network layers.

% VNC challenge
First, accurate knowledge of channel status is hard to be detected at the receiver. 
In recently, cognitive radio mechanisms which interleave channel accesses motivate the frequency
band selection problem of finding the optimal spectrum on which to 
transmit~\cite{ghasemi2008spectrum}. Prior work has considered a number of challenges in
leveraging white space frequencies including spectrum sensing, frequency-agile operation,
geolocation, solving stringent spectral mask requirements, and providing reliable service
in unlicensed and dynamically changing spectrum~\cite{shellhammer2009technical}. 
In particular, there has recently been an acceleration
in spectrum sensing work~\cite{rayanchu2011fluid, kim1996pulse,cabric2004implementation}. Based on 
these works, protocols have been built for multi-channel and/or multiband wireless operation~\cite{MOAR,
raychaudhuri2003spectrum,sabharwal2007opportunistic}.  Other works have presented methods for searching for the most efficient 
transmission channel~\cite{mo2005comparison}, discovering channel information~\cite{rayanchu2011fluid, sabharwal2007opportunistic}, and estimating 
channel quality~\cite{MOAR}.
Moreover, the emergence of a number of diverse sensors on a vehicle motivates work
on heterogeneous wireless networks, which have different frequency bands and
technologies~\cite{hossain2010vehicular}. Thus, the various communication 
standards have diverse throughput capacity, allowing the choice of technology 
to possibly usurp frequency band decisions. For example, an 802.11n link at 5.8 
GHz with high levels of loss
might still be a better choice than a Bluetooth link at 2.4 GHz with little loss
due to the discrepancy of hundreds of Mbps in throughput capacity.
While these works have considered spectral activity and developing protocols and algorithms to 
find spectral holes, less of a focus has been on coupling such information with historical performance in a given 
propagation environment.


Second, in access layer network deployment, specific to rural areas, 
the lack of user density and corresponding traffic
demand per unit area as compared to dense urban areas allows greater levels of
spatial aggregation to reduce the total number of required access points, lowering
network deployment costs. In densely populated urban areas, the greater concentration
of users and higher levels of traffic demand can be served by maximizing the spatial
reuse. While many works have worked to address multihop wireless network deployment
in terms of maximizing served user demand and/or minimizing network costs,
the unique propagation characteristics and the interference from coexisting
activities in white space bands have either not been jointly studied or assumed to 
have certain characteristics without explicit measurement~\cite{si2010overview}. 
Specifically, previous work has investigated wireless 
network deployment in terms of gateway placement, channel assignment, and 
routing~\cite{he2008optimizing,marina2010topology}.
However, each of these works focus on the deployment in WiFi bands without
considering the white space bands. Moreover, the assumption of idle channels
held in these models fails to match the in-field spectrum utility,
which could degrade the performance of a wireless network. These
two issues are critical for designing an optimal network deployment and
providing commercial wireless services to clients in any location.


Third, in backhual layer, researchers have done both centralized and distributed
work for channel assignment and routing~\cite{raniwala2004centralized,wu2006distributed}.
However, FCC's new policy import new diversity to this issue, propagatation
difference becomes a new parameter for the optimization. Since the origin problem
is already a NP-hard problem, the problem developing from the new white space 
policy make it more complex. Thus, new channel assignment methods for the multiband
scenario is a new question people have to answer.

Lastly, beamforming makes the wireless network similar to the wired network
in some degree. However, lacking of protocols and assignment methods make it 
hard to apply the technology in field. The issues of beamforming in application,
such as deafth, new hidden problem need to be leveraged and resolved by new
protocols and algorithms. This new technology inspire the optimization more
challenging and meaningful.


% Contributions of previous work and future work plan
\section{Contributions and Futurework}

This work propose spectrum adaptation for both the link communication and network 
optimization to improve the performance of wireless devices in multiple bands 
accessiable environment.

% VNC Contributions  fixme
In this link spectrum adaptation, we develop multiband adaptation protocols 
which couple the prior knowledge of in-situ performance of various bands with the instantaneous knowledge of 
spectral activity, SNR, and current location of each band to arrive at a decision on the optimal band to transmit. 
To do so, we use an off-the-shelf platform that allows direct comparison and simultaneous experimentation across four different wireless
frequency bands from 450 MHz to 5.8 GHz with the same physical and media access layers. 
We first develop a framework for multiband adaptation using both historical information and 
instantaneous measurements. This framework is broad enough to study adaptation across 
licensed and unlicensed bands, including white space frequency bands.  
We propose two different machine-learning-based multiband adaptation algorithms. The 
first machine learning algorithm, referred to as the Location-based 
Look-up Algorithm, 
is based on the idea of $k$-nearest-neighbor classification. 
The second machine-learning-based algorithm uses decision trees for classification. 
For comparison, we also create two baseline adaptation algorithms which attempt to 
make the optimal band selection based on only: (i.)~historical 
performance data, and (ii.)~instantaneous SNR measurements across 
various bands. 
We perform extensive outdoor V-2-V experiments to evaluate the proposed algorithms.
Our results indicate that the proposed machine learning based algorithms improve
throughput by up to $49.3\%$ over these baseline methods.

% WINMEE

In the access layer network deployment, we perform a measurement study which considers the propagation 
characteristics and observed in-field spectrum availability of white space
and WiFi channels to find the total number of access points required to serve a 
given user demand. Across varying population densities in representative 
rural and metropolitan areas, we compare the cost savings (defined in terms of
number of access points reduced) when white space bands are not used.
To do so, we first define the metric to quantify the spectrum utility in a
given measurement location. With the in-field measured spectrum utility data 
in metropolitan and surrounding areas of Dallas-Fort Worth (DFW), we 
calculate the activity level in WiFi and white space bands. Second, we 
propose a measurement-driven framework to find the number of access points required 
for areas with differing population densities according to our measurement locations
and census data. We then evaluate our measurement-driven framework, showing
the band selection across downtown, residential and university settings in
urban and rural areas and analyze the impact of white space and WiFi
channel combinations on a wireless deployment in these representative scenarios.
% Paper contributions
The main contributions of our work are as follows:
We perform in-field measurements of spectrum utilization in various representative
scenarios across the DFW metroplex, ranging from sparse rural to dense urban areas and 
consider the environmental setting (e.g., downtown, residential, or university campus).
We develop a measurement-driven Multi-band Access Point Estimation (MAPE) framework 
to jointly leverage propagation and spectrum availability of white space and WiFi bands 
for wireless access networks across settings.
We analyze our framework under capacity and coverage constraints 
to show that, with white space bands, the number of access points can be greatly
reduced from WiFi-only deployments by up to 1650\% in rural areas.
We quantify the impact of white space and WiFi channel
combinations to understand the tradeoffs involved in choosing the optimal channel setting,
given a certain number of available channels from multiple bands.
Furthermore, we discuss to centralized using frequencies or distributed using them in 
single access point. Through our numerical simulation, we show that in dense area
centralized using spectrum gain more than sparse area.


In the backhual layer network deployment, we leverage the diversity in propagation of
white space and WiFi bands in the planning and deployment
of large-scale wireless mesh networks. To do so, we first form an
integer linear program to jointly exploit 
white space and WiFi bands for optimal WhiteMesh topologies.
Second, since similar problem formulations have been shown to be 
NP-hard~\cite{jain2005impact}, we design a heuristic algorithm, 
Band-based Path Selection (BPS). We then
show the algorithm approaches the performance of the 
optimal solution but with a reduced complexity. To assess the 
performance of our scheme, we compare the performance of 
BPS against two well-known multi-channel, multi-radio deployment 
algorithms across a wide range of scenarios, including those
typical for rural areas as well as urban settings. Finally, we
quantify the degree to which the joint use of both band types can improve the 
performance of wireless mesh networks.
We develop an optimization framework based on integer
linear programming to jointly leverage white space and WiFi bands
to serve the greatest user demand in terms of gateway throughput 
in wireless mesh networks.  
We build an algorithm, Band-based Path 
Selection (BPS), which considers the diverse propagation 
and overall interference level of WiFi and white space bands using
a two-stage approach.  In the first stage, we prioritize the bands
with the greatest propagation to reduce the overall hop count. In the
second stage, we compare the interference level of path choices with
similar hop count. 
We perform extensive analysis across offered loads,
network sizes, and WiFi/white space band combinations, showing that BPS outperforms existing
multi-channel, multi-radio algorithms techniques by 3 and 6 times
in terms of the served user demand.  
Given similar channel resources
(bandwidth and transmission power), we additionally show that while
WiFi-only mesh topologies would largely outperform mesh networks with only
white space bands, the joint use of the two types of bands (i.e., WhiteMesh 
networks) can yield up to 170\% of the served user demand compared to mesh networks
with only one type of band.



In the beamforming application, we are going to design protocols to 
optimize the network deployment cost under the limitation of QoS. Through
the beamforming, we are going to increase the time reuse in the traditional
RTS/CTS link communication and the reduce the cost for network deployment.
  


% THesis Overview
\section{Proposal Overview}

In Chapter~\ref{ch:background}, we discuss the background of this work, including the basic
knowledge related to this work and the hardware platform and software tools that we 
use. 
Chapter~\ref{ch:vnc} provides an analysis and solution of vehicular link spectrum adaptation
for multiband scenario. We also proposed an access layer network deployment framework
in Chapter~\ref{ch:winmee} and discuss the optimization of a single access point with 
spectrum limitation.
In Chapter~\ref{ch:wm}, we propose an algorithm to assign channels in multiband scenario
for optimizing the throughput performance in backhual networks.
In Chapter~\ref{ch:futurework}, we discuss the potential beamforming technology application
in link communication and networks.
Finally, we conclude our work in Chapter~\ref{ch:conclusion}.

